\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[colorlinks=true,linkcolor=blue!60!black,urlcolor=blue!60!black,citecolor=blue!50!black]{hyperref}
\usepackage{cleveref}

\newcommand{\Jpm}{J_{\pm}}
\newcommand{\Jzz}{J_{zz}}
\newcommand{\ghex}{g_{\mathrm{hex}}}
\newcommand{\gfour}{g_{4}}
\newcommand{\Heff}{H_{\mathrm{eff}}}
\newcommand{\Tpk}{T_{\mathrm{peak}}}
\newcommand{\bvarphi}{\bm{\varphi}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\bN}{\mathbf{N}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bL}{\mathbf{L}}
\newcommand{\bq}{\mathbf{q}}
\newcommand{\brho}{\bm{\rho}}
\newcommand{\bdelta}{\bm{\delta}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Ds}{\Delta_{\mathrm{s}}}
\newcommand{\twoinout}{2\text{-in-}2\text{-out}}
\newcommand{\Pice}{\mathcal{P}}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\title{Unbiased exact diagonalization of quantum spin ice in the
$\pi$-flux regime:\\
a zero-transport projector for finite-size gauge contamination}
\author{}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
The thermodynamics and dynamics of quantum spin ice (QSI) are governed
by an emergent compact $U(1)$ gauge field. Its unfrustrated
\emph{$0$-flux} sector ($\Jpm>0$) is accessible to sign-problem-free
quantum Monte Carlo (QMC); the frustrated \emph{$\pi$-flux} sector
($\Jpm<0$) realized in the dipolar--octupolar cerium pyrochlores carries
a QMC sign problem, leaving exact diagonalization (ED) as the only
unbiased finite-temperature method. ED, however, has a finite-size
problem specific to QSI: a periodic cluster hosts short ice-preserving
ring exchanges that wind the torus and have no infinite-lattice
analogue. The dominant ones are four-site loops with coupling
$\gfour=4\Jpm^2/\Jzz$, which is \emph{second} order in $\Jpm$ and
therefore parametrically larger than the genuine photon coupling
$\ghex=12|\Jpm|^3/\Jzz^2$; they contaminate exactly the emergent-gauge
observables one wants (the low-$T$ specific-heat peak and the low-energy
$S^{zz}(\bq,\omega)$). We remove this contamination with a
\emph{zero-transport projector}: retain only ice-manifold processes
whose net emergent-charge transport $\bdelta=\brho+\bN\!\cdot\!\bL$
vanishes modulo the cluster periods. We prove, both mathematically and
by direct computation, that the projector (i)~annihilates every winding
ring exchange at all orders in $\Jpm$, (ii)~preserves every contractible
ring exchange at full strength, and (iii)~coincides with the
operator-level twist average taken in the translationally-invariant
gauge -- clarifying why the naive twist averages used in practice fail
(the boundary gauge removes only half; averaging observables rather than
the Hamiltonian keeps the full spurious scale, gauge-invariantly). On
the $16$- and $32$-site clusters the projector moves the low-$T$
specific-heat peak and the low-energy $S^{zz}$ weight from the artifact
scale $\gfour$ to the photon scale $\ghex$, for both flux signs. This
makes a $32$-site pyrochlore an unbiased laboratory for the $\pi$-flux
gauge sector, to be benchmarked against $0$-flux QMC and applied at the
exchange parameters of the cerium pyrochlores.
\end{abstract}

%=====================================================================
\section{The problem}
\label{sec:problem}

\paragraph{Quantum spin ice and its emergent photon.}
The minimal model of QSI is the XXZ antiferromagnet on the pyrochlore
lattice, written in the local $\langle111\rangle$ frames,
\begin{equation}
H=\Jzz\sum_{\langle ij\rangle}S^z_iS^z_j
-\Jpm\sum_{\langle ij\rangle}\bigl(S^+_iS^-_j+S^-_iS^+_j\bigr),
\qquad 0<|\Jpm|\ll\Jzz .
\label{eq:H}
\end{equation}
For $\Jpm\to0$ the ground manifold is the exponentially degenerate set
of $\twoinout$ (ice-rule) states. A small $\Jpm$ lifts this degeneracy
through virtual ring exchanges and generates the deconfined $U(1)$
lattice gauge theory of QSI: $S^z$ is the emergent electric field, the
third-order hexagonal ring exchange
\begin{equation}
\ghex=\frac{12\,|\Jpm|^3}{\Jzz^2}
\label{eq:ghex}
\end{equation}
is the Maxwell term whose quantum $=$ the emergent photon, and ice-rule
defects are gapped, fractionalized spinons (pair threshold
$\Ds\approx0.8\,\Jzz$ on the clusters below). The \emph{sign} of the
ring coupling follows $\mathrm{sgn}(\Jpm^3)$: $\Jpm>0$ gives the
unfrustrated $0$-flux vacuum, $\Jpm<0$ the frustrated $\pi$-flux vacuum.
Thermodynamically the photon appears as a low-temperature Schottky peak
in $C(T)$ at $T\sim\ghex$, far below the spinon (charge) peak at
$T\sim\Jzz$; dynamically it appears as low-energy spectral weight in the
non-spin-flip channel $S^{zz}(\bq,\omega)$.

\paragraph{Why $\pi$-flux needs exact diagonalization.}
In the $S^z$ basis the off-diagonal matrix elements of \cref{eq:H} are
$-\Jpm$. For $\Jpm>0$ they are negative and worldline/stochastic-series
QMC is sign-problem-free: the $0$-flux QSI is quantitatively
characterized. For $\Jpm<0$ the same elements are positive and QMC
acquires a sign problem. The $\pi$-flux regime -- believed to describe
the cerium pyrochlores Ce$_2$Zr$_2$O$_7$, Ce$_2$Sn$_2$O$_7$,
Ce$_2$Hf$_2$O$_7$ -- is thus a QMC dead zone, and unbiased
finite-temperature information must come from exact diagonalization of a
finite cluster.

\paragraph{The finite-size problem.}
A periodic cluster is a flat torus, and its periodic identifications
generate ice-preserving ring exchanges that \emph{wind} the torus and do
not exist on the infinite lattice. The shortest and most damaging are
\emph{four-site} loops. Because a four-site loop factorizes into two
disjoint nearest-neighbour (NN) hops, it is generated at \emph{second}
order in $\Jpm$,
\begin{equation}
\gfour=\frac{4\,\Jpm^2}{\Jzz},
\qquad\text{so}\qquad
\frac{\gfour}{\ghex}=\frac{\Jzz}{3|\Jpm|}\gg1 .
\label{eq:g4}
\end{equation}
The artifact is thus \emph{parametrically larger} than the physics, and
it grows as $|\Jpm|\to0$ -- exactly the weak-coupling regime of the
cerium materials. It lives entirely in the ice manifold, so it
contaminates precisely the emergent-gauge observables: it pushes the
low-$T$ $C(T)$ peak and the low-energy $S^{zz}$ weight up to the scale
$\gfour$ instead of $\ghex$ (quantified in \cref{sec:demo}). The spinon
continuum ($S^{+-}$ at $\omega\sim\Jzz$) lies outside the ice manifold
and is \emph{not} contaminated.

\paragraph{Goal.}
Remove this finite-size contamination from the gauge sector of a
periodic-cluster ED, rigorously and verifiably, so that the $32$-site
pyrochlore delivers unbiased $\pi$-flux thermodynamics and dynamical
structure factor.

%=====================================================================
\section{The effective gauge Hamiltonian and the winding loops}
\label{sec:setup}

\paragraph{Cluster torus.}
The cluster carries $N$ sites with periodic identifications generated by
the rows of an integer matrix $\bL\in\Z^{3\times3}$ (in units of the
cubic lattice constant $a$); position space is the flat torus
$T^3=\mathbb R^3/\bL\Z^3$ with first homology $\Z^3$. Each NN bond
$(i,j)$ has a \emph{wrap vector}
$\bn_{ij}=\mathrm{round}[(\br_j-\br_i)\bL^{-1}]$, zero in the interior
and nonzero across the boundary.

\paragraph{Effective Hamiltonian.}
Degenerate ($\Jpm$-)perturbation theory in the ice manifold yields
$\Heff=\sum_{\mathcal C} c_{\mathcal C}\,O_{\mathcal C}$, a sum of
ring-exchange operators
$O_{\mathcal C}=(S^+_aS^-_bS^+_cS^-_d\cdots)$ over ice-preserving cycles
$\mathcal C$; because each intermediate resolvent is Ising
($\propto\Jzz$), the coefficient $c_{\mathcal C}$ is exactly
$\propto\Jpm^{|\mathcal C|/2}$. On the $2{\times}2{\times}2$ FCC cluster
($N=32$) the ice manifold has dimension $2970$, so $\Heff$ is
\emph{fully diagonalizable} and its thermodynamics is numerically exact.
Enumeration (Schrieffer--Wolff with per-process bookkeeping) gives the
census in \cref{tab:census}: the winding four-loops and wrapping
hexagons are the contamination; the contractible hexagons carry the
photon.

\begin{table}[t]
\centering
\caption{Census of ice-preserving ring exchanges on the two production
clusters. ``winding/wrapping'' loops close only through the periodic
boundary ($\bw\neq\bm0$); ``contractible'' hexagons are genuine bulk
plaquettes ($\bw=\bm0$). The last column is the survival fraction under
the zero-transport projector of \cref{sec:projector}.}
\label{tab:census}
\begin{tabular}{llrcc}
\toprule
cluster & channel & count & order & kept by projector \\
\midrule
\multirow{3}{*}{$1{\times}1{\times}1$ cubic ($N{=}16$)}
 & winding four-loops   & 36 & $\Jpm^2$ & $0$ \\
 & wrapping hexagons    & 48 & $\Jpm^3$ & $0$ \\
 & contractible hexagons& 16 & $\Jpm^3$ & $1$ \\
\midrule
\multirow{3}{*}{$2{\times}2{\times}2$ FCC ($N{=}32$)}
 & winding four-loops   & 24 & $\Jpm^2$ & $0$ \\
 & wrapping hexagons    & 96 & $\Jpm^3$ & $0$ \\
 & contractible hexagons& 32 & $\Jpm^3$ & $1$ \\
\bottomrule
\end{tabular}
\end{table}

%=====================================================================
\section{The zero-transport projector}
\label{sec:projector}

The gauge-invariant label that distinguishes a genuine (bulk) ring
exchange from a torus-winding one is not homology but \emph{charge
transport}.

\begin{definition}[Net charge transport]
For an ice-manifold process taking configuration $c$ to $c'$ by raising
a site set $R$ and lowering a site set $L$ through hops of total wrap
$\bN=\sum\bn$, the net transport is
\begin{equation}
\bdelta=\brho+\bN\!\cdot\!\bL,
\qquad
\brho=\sum_{i\in R}\br_i-\sum_{j\in L}\br_j ,
\label{eq:delta}
\end{equation}
the physical displacement of the transported $S^z$ (boson) density.
$\brho$ is the intrinsic dipole of the flipped-site set (home
positions); $\bN\!\cdot\!\bL=\sum_k N_k\bL_k$ is the wrap contribution.
\end{definition}

\begin{definition}[Zero-transport projector]
$\Pice$ retains exactly the processes with $\bdelta\equiv\bm0
\pmod{\bL}$, i.e.\ $\bdelta\in\bL\Z^3$, and deletes the rest.
\end{definition}

Two implementations, cross-validated in \cref{sec:proof-comp}:
\emph{(i)} in perturbation theory, a row filter on the enumerated
processes; \emph{(ii)} nonperturbatively, a mask on the exactly
downfolded ice-manifold Hamiltonian, keeping matrix element $(c,c')$ iff
its flip-set dipole $\brho\equiv\bm0\pmod\bL$. Because
$\bN\!\cdot\!\bL\in\bL\Z^3$ automatically, the criterion
$\bdelta\in\bL\Z^3$ reduces to $\brho\in\bL\Z^3$: \emph{the projector's
action on an operator depends only on its dipole, at every order.}

%=====================================================================
\section{Proof that the projector works: mathematics}
\label{sec:proof-math}

\subsection{Ring exchanges are sums over virtual paths}

The naive expectation is that a winding ring exchange picks up its
homology holonomy $e^{-i\bw\cdot\bvarphi}$ under a boundary twist
$\bvarphi$ and so is removed by averaging over $\bvarphi$. This is
false, and the reason underlies everything.

\begin{theorem}[Matched-pair factorization]
\label{thm:matched}
Let $\mathcal C$ be an ice-preserving simple cycle of length $2m$ with
winding $\bw$. At the leading order at which $O_{\mathcal C}$ appears,
its coefficient in the boundary-gauge twisted $\Heff(\bvarphi)$ is
\begin{equation}
c_{\mathcal C}(\bvarphi)=-\frac{g_{\mathcal C}}{M}\sum_{\mu=1}^{M}
e^{-i\bN_\mu\cdot\bvarphi},
\qquad \sum_\mu\bN_\mu=\bw \ (M{=}2),
\label{eq:matched}
\end{equation}
where $\mu$ runs over the $M\ge2$ perfect matchings of $\mathcal C$ into
disjoint NN dimer moves and $\bN_\mu$ is the total wrap of matching
$\mu$. For a two-matching loop the gauge-invariant magnitude is
\begin{equation}
|c_{\mathcal C}(\bvarphi)|=g_{\mathcal C}\bigl|\cos(\tfrac12\bw\cdot\bvarphi)\bigr|.
\label{eq:mag}
\end{equation}
\end{theorem}

\begin{proof}
At leading order $O_{\mathcal C}$ is generated only by flipping each of
its $2m$ spins once, i.e.\ by a product of $m$ disjoint NN dimer moves --
a perfect matching of the cycle; an even cycle has $M\ge2$. Each hop
$S^+_iS^-_j$ carries $e^{-i\bn_{ij}\cdot\bvarphi}$, so matching $\mu$
carries $e^{-i\bN_\mu\cdot\bvarphi}$ with $\bN_\mu=\sum_{\text{hops}}\bn$.
The $M$ matchings share amplitude $g_{\mathcal C}/M$ (equal denominators
and weight). For a $4$-cycle $abcd$ the two matchings $\{ab,cd\}$ and
$\{bc,da\}$ together traverse each bond once, so
$\bN_1+\bN_2=\bw$; \eqref{eq:mag} follows from
$|1+e^{-i\bw\cdot\bvarphi}|=2|\cos(\tfrac12\bw\cdot\bvarphi)|$.
\end{proof}

The operator thus lives in \emph{several} transport sectors at once; the
holonomy is only the \emph{relative} phase of the matchings, and the
magnitude \eqref{eq:mag} \emph{vanishes} at $\bw\cdot\bvarphi=\pi$ --
which a single homology phase (constant magnitude) never does.

\subsection{The projector removes winding, keeps contractible}

\begin{theorem}[Transport selection]
\label{thm:transport}
Let $\ell_0=\min_{\bw\neq\bm0}|\bw\cdot\bL|$ be the shortest nonzero
cluster period. If every ring exchange under consideration has
alternating dipole $|\brho_{\mathcal C}|<\ell_0$, then $\Pice$ deletes
every loop with $\brho_{\mathcal C}\neq\bm0$ at all orders in $\Jpm$ and
keeps every loop with $\brho_{\mathcal C}=\bm0$ at full strength; on the
production clusters these are exactly the winding and the contractible
loops respectively.
\end{theorem}

\begin{proof}
By the remark below \cref{eq:delta}, $\Pice$ keeps $O_{\mathcal C}$ iff
$\brho_{\mathcal C}\in\bL\Z^3$, at every order. Since
$|\brho_{\mathcal C}|<\ell_0$, the only lattice vector it can equal is
$\bm0$: hence $\brho_{\mathcal C}\in\bL\Z^3\iff\brho_{\mathcal C}=\bm0$,
and $\Pice$ keeps $O_{\mathcal C}$ iff $\brho_{\mathcal C}=\bm0$.

A genuine bulk plaquette (regular $[111]$-hexagon, or its
four-tetrahedron analogue) is centrosymmetric -- its raised and lowered
sites share a centroid -- so $\brho=\bm0$ and it is kept. A winding loop
has $\brho\neq\bm0$: were $\brho=\bm0$, then in the two-matching
factorization $\bdelta_1+\bdelta_2=\bw\!\cdot\!\bL$ with
$|\bw\!\cdot\!\bL|\ge\ell_0$ would force some
$|\bdelta_\mu|\ge\tfrac12\ell_0$, yet $\bdelta_\mu=\brho+\bN_\mu\bL$ with
$\brho=\bm0$ makes $\bdelta_\mu$ a lattice vector, hence $\bm0$ or
$\ge\ell_0$ -- contradiction. So winding loops are deleted.
\end{proof}

On the production clusters the hypothesis holds with room to spare: the
alternating dipole of \emph{every} four-loop and wrapping hexagon has the
same magnitude $|\brho|=a/\sqrt2$, below $\ell_0=a$ (cubic) and
$\sqrt2\,a$ (FCC), while every contractible hexagon has $\brho=\bm0$
identically -- giving the $0/0/1$ survival pattern of \cref{tab:census}.
The single residual bias
is honest: a \emph{contractible} process with a nonzero dipole
($\brho\neq\bm0$, possible only for non-centrosymmetric loops at order
$\Jpm^4$ and beyond) would also be removed; within the NN model these are
$O(\Jpm^4/\Jzz^3)$, parametrically below $\ghex$.

\subsection{Relation to twist averaging: why the naive versions fail}

\begin{theorem}[Twist-average identity]
\label{thm:twist}
In the translationally-invariant (``smooth'') gauge -- a uniform vector
potential $\mathbf A=\bL^{-1}\bvarphi$ placing
$\theta_{ij}=\mathbf A\cdot\mathbf d_{ij}$ on \emph{every} bond -- each
process carries $e^{-i\mathbf A\cdot\bdelta}$, and the operator-level
twist average over the extended period is $\Pice$:
\begin{equation}
\overline{\Heff}^{\,\rm smooth}
=\int \Heff(\bvarphi)\,\frac{d\bvarphi}{(\text{period})}
=\Pice\,\Heff .
\end{equation}
\end{theorem}

\begin{proof}
A hop $i\to j$ contributes $\mathbf A\cdot\mathbf d_{ij}$; summed along a
process the phase is $\mathbf A\cdot(\sum\mathbf d)=\mathbf A\cdot\bdelta$
with $\bdelta$ its transport. Averaging $e^{-i\mathbf A\cdot\bdelta}
=e^{-i\bvarphi\cdot(\bL^{-1}\bdelta)}$ over the period on which the
smooth-gauge $\Heff$ is single-valued gives $\delta_{\bL^{-1}\bdelta\in
\Z^3}=\delta_{\bdelta\in\bL\Z^3}$, i.e.\ $\Pice$.
\end{proof}

So the projector \emph{is} twist averaging, done correctly. This pins
down the two ways the naive procedure fails, both verified numerically
in \cref{sec:proof-comp}:
\begin{corollary}[Failure modes]
\label{cor:fail}
\emph{(a) Wrong gauge.} In the boundary-bond gauge, a winding loop's
zero-wrap matching ($\bN_\mu=\bm0$) is phase-free, so the operator
average keeps $\tfrac12g_{\mathcal C}$ rather than $0$; the ``half'' is a
gauge artifact.
\emph{(b) Wrong object.} Averaging an \emph{observable} $\mathcal O(g)$
(a specific $C(T)$ or $S(\bq,\omega)$) rather than the Hamiltonian keeps
every \emph{even} power of the gauge-invariant $|c_{\mathcal C}(\bvarphi)|
=g_{\mathcal C}|\cos(\tfrac12\bw\cdot\bvarphi)|$, because
$\overline{|c|^{2k}}\neq0$; the spurious scale survives in \emph{any}
gauge. Since practice averages the measured quantity, naive twist
averaging fails.
\end{corollary}

%=====================================================================
\section{Proof that the projector works: computation}
\label{sec:proof-comp}

Each mathematical claim is checked exactly on the $16$-site cluster
(where full ED is available) and, where relevant, on the $32$-site FCC
cluster.

\paragraph{(1) Matched-pair magnitude \eqref{eq:mag}.}
For a winding four-loop [sites $(0,2,8,10)$, $\bw=(-1,0,-1)$] the exact
second-order coefficient obeys, along $\bw$,
\[
|c(\bvarphi)|/\gfour = 1.000,\,0.866,\,0.707,\,0.500,\,0.000,\dots
= |\cos(\tfrac12\bw\cdot\bvarphi)|\ \text{to all digits,}
\]
vanishing at $\bw\cdot\bvarphi=\pi$. A single homology phase would keep
$|c|=\gfour$ for all $\bvarphi$; the vanishing is the two-matching
interference of \cref{thm:matched}.

\paragraph{(2) Transport selection \eqref{thm:transport}.}
Direct enumeration confirms \cref{tab:census}: on both clusters
\emph{every} winding four-loop and wrapping hexagon has
$\brho\notin\bL\Z^3$ (deleted), while \emph{every} contractible hexagon
has $\brho=\bm0$ (kept) -- survival $0/0/1$, to machine precision, at
orders $\Jpm^2$ and $\Jpm^3$.

\paragraph{(3) Twist-average identity \eqref{thm:twist} and failure modes.}
The coefficient of the winding four-loop under the operator-level twist
average is
\[
-\gfour\ (\text{bare})\ \longrightarrow\
-\tfrac12\gfour\ (\text{boundary gauge})\ \longrightarrow\
0\ (\text{smooth gauge}=\Pice),
\]
while a contractible hexagon is $+\ghex$ in all three -- confirming both
\cref{thm:twist} and \cref{cor:fail}(a). For \cref{cor:fail}(b), the
corner-averaged \emph{observable} keeps the spurious scale
(\cref{sec:demo}).

\paragraph{(4) Nonperturbative agreement.}
The dipole mask on the exactly downfolded $16$-site Hamiltonian
reproduces the perturbative $\Pice$: e.g.\ the low-$T$ $C(T)$ peak at
$\Jpm=-0.05$ agrees to $0.00122$ vs.\ $0.00127\,\Jzz$, validating the
truncation.

%=====================================================================
\section{Demonstration: the artifact removed, both flux sectors}
\label{sec:demo}

\paragraph{Thermodynamics.}
\Cref{tab:thermo} lists the low-$T$ (gauge) peak of $C(T)$ in units of
$\ghex$. Bare periodic ED places it at the artifact scale, drifting up
as $\gfour/\ghex=\Jzz/3|\Jpm|$ toward weak coupling; the projector
brings it to $O(\ghex)$ for both flux signs. Corner-averaging the
\emph{observable} barely moves it (\cref{cor:fail}(b)).

\begin{table}[t]
\centering
\caption{Low-$T$ specific-heat (gauge) peak $\Tpk/\ghex$ on the
$32$-site FCC cluster (exact full spectrum of the $2970$-state ice
manifold, order $\Jpm^4$). Bare periodic ED sits at the artifact scale;
$\Pice$ restores the photon scale $O(1)$.}
\label{tab:thermo}
\begin{tabular}{lccc}
\toprule
$\Jpm/\Jzz$ & bare PBC & $\Pice$ & $\gfour/\ghex$ \\
\midrule
$-0.03$ & $5.4$ & $0.6$ & $11.1$ \\
$-0.05$ & $3.1$ & $0.3$ & $6.7$ \\
$+0.046$ (0-flux) & $10.1$ & $0.8$ & $7.2$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Dynamical structure factor.}
The emergent photon is the low-energy weight of the non-spin-flip
channel $S^{zz}(\bq,\omega)=\sum_n|\langle n|S^z_\bq|0\rangle|^2
\delta(\omega-\omega_n)$; since $S^z$ is diagonal in the ice basis this
is computed exactly from the projected $\Heff$. \Cref{tab:dssf} gives
the low-$\omega$ weight centroid: bare ED places it at the artifact
scale $\gfour$, the observable twist-average leaves it there, and only
$\Pice$ brings it to the photon scale. (The spinon continuum $S^{+-}$ at
$\omega\sim\Jzz$ is uncontaminated and is obtained separately by
full-Hilbert-space methods.)

\begin{table}[t]
\centering
\caption{Low-$\omega$ centroid of the emergent-photon structure factor
$S^{zz}$ (in units of $\gfour$) at $\Jpm=-0.05\,\Jzz$. Twist averaging
of the observable fails; the projector reaches the photon scale
$\sim\ghex=0.15\,\gfour$.}
\label{tab:dssf}
\begin{tabular}{lcc}
\toprule
scheme & $16$-site & $32$-site FCC \\
\midrule
bare PBC                         & $4.2\,\gfour$ & $1.8\,\gfour$ \\
twist-averaged observable (8 corners) & $2.6\,\gfour$ & -- \\
zero-transport projector $\Pice$ & $0.47\,\gfour$ & $0.28\,\gfour$ \\
\bottomrule
\end{tabular}
\end{table}

Both tables make the same point in the two experimentally relevant
observables: the finite-size artifact sits at $\gfour$, twist averaging
does not remove it, and $\Pice$ restores the photon scale $\ghex$ for
both $0$- and $\pi$-flux.

%=====================================================================
\section{Outlook: the $\pi$-flux campaign}
\label{sec:outlook}

The projector makes the $32$-site pyrochlore an unbiased calculator of
the QSI gauge sector, and the couplings scale as $\Jpm^k$ exactly, so a
single effective-Hamiltonian build serves every coupling and both signs.
The program is: \emph{(A)} benchmark the projected $0$-flux $C(T)$,
$S(T)$ and photon $S^{zz}$ against sign-problem-free QMC at a common
coupling; \emph{(B)} cross into $\pi$-flux and compute the same
quantities where QMC cannot, resolving the emergent-photon lower peak;
\emph{(C)} evaluate at the measured exchange parameters of
Ce$_2$Zr$_2$O$_7$, Ce$_2$Sn$_2$O$_7$ and Ce$_2$Hf$_2$O$_7$ and confront
the neutron $S(\bq,\omega)$ and the specific heat directly. The
uncontaminated spinon continuum is supplied by full-Hilbert-space
finite-temperature Lanczos, with twist averaging used for its legitimate
purpose there -- Brillouin-zone interpolation of the genuine bands, not
contamination removal.

\end{document}
